Effectful Semantics for Substructural SSA

### 2023-11-05 pldi24/paper.tex
Tao

\TODO{relate separation logic trick to \cite{promonad}, \cite{linear-state-usage}}

@@

\begin{definition}[Elgot category] 
  An \textbf{Elgot category} is a category \(\mc{C}\) with coproducts equipped with an operator \(\cdot^\dagger: \mc{C}(A, B + A) \to \mc{C}(A, B)\) satisfying the following properties:
  \begin{itemize}
    \item \textbf{Fixpoint:} \(f;[\ms{id}, f^\dagger] = f^\dagger\)
    \item \textbf{Naturality:} \((f;g + \ms{id})^\dagger = f^\dagger;g\)
    \item \textbf{Codiagonal:} \(\forall f \in \mc{C}(A, (B + A) + A), (f^\dagger)^\dagger = (f;[\ms{id}, \iota_1])^\dagger\)
    \item \textbf{Uniformity:}
    \(
      h;f = g;\ms{id} + h \implies h;f^\dagger = g^\dagger
    \)
  \end{itemize}
  A monad \(\mb{T}\) on \(\mc{C}\) such that its Kleisli cateogry \(\mc{C}_{\mb{T}}\) is Elgot is called an \textbf{Elgot monad}.
\end{definition}
\TODO{why do these matter? they show up in the proofs, but...}
Some important properties of all Elgot categories include:
\begin{itemize}
  \item \textbf{Dinaturality:} \((g;[\iota_0, h])^\dagger = g;[\ms{id}, (h;[\iota_0, g])^\dagger]\). \TODO{Proposition 16 of \cite{goncharov-guarded-unguarded}; also Lemma 31 of \cite{goncharov-squaring}}
  \item \textbf{Squaring:} \((f;[\iota_0, f])^\dagger = f^\dagger\). \TODO{Lemma 31 of \cite{goncharov-squaring}}
\end{itemize}
\TODO{Hyland and Hasegawa traces?}
\TODO{intuition? Can draw diagrams, but those take up space}

@@

\TODO{example of Elgot monads: traces}

>>>>

\TODO{relate separation logic trick to \cite{promonad}, \cite{linear-state-usage}, \cite{mellies-ftrs}}

@@

\begin{definition}[Elgot category] 
  An \textbf{Elgot category} is a category \(\mc{C}\) with coproducts equipped with an operator \(\cdot^\dagger: \mc{C}(A, B + A) \to \mc{C}(A, B)\) satisfying the following properties:
  \begin{itemize}
    \item \textbf{Fixpoint:} \(f;[\ms{id}, f^\dagger] = f^\dagger\)
    \item \textbf{Naturality:} \((f;g + \ms{id})^\dagger = f^\dagger;g\)
    \item \textbf{Codiagonal:} \(\forall f \in \mc{C}(A, (B + A) + A), (f^\dagger)^\dagger = (f;[\ms{id}, \iota_1])^\dagger\)
    \item \textbf{Uniformity:}
    \(
      h;f = g;\ms{id} + h \implies h;f^\dagger = g^\dagger
    \)
  \end{itemize}
  A monad \(\mb{T}\) on \(\mc{C}\) such that its Kleisli cateogry \(\mc{C}_{\mb{T}}\) is Elgot is called an \textbf{Elgot monad}.
\end{definition}
\TODO{why do these matter? they show up in the proofs, but...}
Some important properties which hold for all Elgot categories include:
\begin{itemize}
  \item \textbf{Dinaturality:} \((g;[\iota_0, h])^\dagger = g;[\ms{id}, (h;[\iota_0, g])^\dagger]\)
  \item \textbf{Squaring:} \((f;[\iota_0, f])^\dagger = f^\dagger\)
\end{itemize}
Proofs can be found in Lemma 31 of \cite{goncharov-squaring}.

\TODO{Hyland and Hasegawa traces?}

\TODO{intuition? Can draw diagrams, but those take up space}

@@
